---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF - SS 2022/23"
author: "Simon Herrmann (11908690)"
date: "`r Sys.Date()`"
---
    
**Important information:** Unfortunately, the Git GUI of RStudio didn't work in my case. I spoke with Jan about my problem in the practical and he gave me the advice to use another IDE. I tried VSCode - but unfortunately Git didn't work in VSCode either. So I decided shorthand to use the terminal (zsh) to create the commands for Git. I inserted all my zsh-commands in gray boxes in the solution here. 

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

### Exercise 1a) 

**Task:** Start by initializing a new R project with git support, called
    `exeRcise-sheet-3`. If you forgot how to do this, you can follow
    this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
    
**My approach:** In this case I initialized Git with the RStudio GUI *"New project" > "Checkbox: Create a git repository"*

### Exercise 1b) 

**Task:** Commit the files generated by Rstudio.

````markdown
#change directory
cd "<path to project directory>"

#check existing files
git status

#stage files generated by RStudio
git add .gitignore
git add exeRcise-sheet-3.Rproj

#commit files with comment
git commit -m "Initial files generated by R Studio"
````

### Exercise 1c) 

**Task:** For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc. Note: *This applies only to answers that have text or code as their answer. If you complete tasks in a different order or forget to commit one, this is no problem. If you change your answers you can just create multiple commits to track the changes.*
    
**Information:** Before starting with the next exercises, I copied the Exercise sheet #3 into the project folder. 

### Exercise 1d) 

**Task:** Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)

**Two strengths:**

* Old files that would have been "deleted" without Git can't be restored anytime
* Git facilitates working in bigger teams with other developers

**Two weaknesses:**

* Setting up Git can be complicated (especially topics like SSH-authentication or setting up customized workflows)
* Using Git while programming costs more time than simply going without it

### Exercise 1e) 

**Task:** Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves.
    
**Solution:** First, I added the file *ExerciseSheet_3_Herrmann.html* to the .gitignore-file
    
````markdown
git add .gitignore
git commit -m "Exercise 1e"
````

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

### Exercise 2a) 

**Task:** Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:
    
**Link:** https://github.com/SimonAtLMU/exeRcise-sheet-3

### Exercise 2b) 

**Task:** Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
    
````markdown
git remote add origin git@github.com:SimonAtLMU/exeRcise-sheet-3.git
git branch -M main
git push -u origin main
````

Then, I successfully inserted my passphrase for the private SSH key

### Exercise 2c) 

**Task:** Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.
    

## Exercise 3: Baby-Names in Munich (4.5 Points)

Download the latest open datasets on given names ("Vornamen") from the
open data repository of the city of Munich for the years `2022` and
`2021`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

### Exercise 3a)

**Task:** Download the data for both years and track it in git. For small
    datasets like these adding them to git is not a problem.
    
**Information:** I downloaded the data and inserted them in a new folder called *data*

````markdown
git add data
git commit -m "Added new dataframes"
````

### Exercise 3b)

**Task:** Load the data for both years into R. Check the type of the count
    variable ("Anzahl") and look into the data to determine why it is
    not numeric? Fix the problem in an appropriate manner, it is OK if
    some of the counts are inaccurate because of this. Explain your
    solution and the repercussions.
    
```{r message=FALSE, warning=FALSE}
#assign the downloaded df to objects
names21 <- read.csv(file.path("data", "vornamen_2021.csv"))

names22 <- read.csv(file.path("data", "open_data_portal_2022.csv"))

#check type of "Anzahl"-variables
c(typeof(names21$Anzahl), typeof(names22$Anzahl))

#Make "Anzahl"-variables numeric; solution for "4 oder weniger"
library(tidyverse)

#df for 2021
names21 <- names21 %>% 
  rowwise() %>% 
  mutate(Anzahl = if_else(Anzahl == "4 oder weniger", as.character(runif(1, min = 0, max = 4)), Anzahl)) %>% #generate random values for "4 oder weniger"
  ungroup() %>% 
  mutate(Anzahl = as.numeric(Anzahl))

#df for 2022
names22 <- names22 %>% 
  rowwise() %>% 
  mutate(Anzahl = if_else(Anzahl == "4 oder weniger", as.character(runif(1, min = 0, max = 4)), Anzahl)) %>% #generate random normal-distributed values for "4 oder weniger"
  ungroup() %>% 
  mutate(Anzahl = as.numeric(Anzahl))

```
    
**Explanation of solution:** First of all, I realised that the values in the column "Anzahl" have the value of *character*. Based on the task explanation, I assumed that these values should be converted to a numeric type. Since the values "4 oder weniger" could be everything between 0 and 4, I replaced these values with random values between 0 and 4 (specified with the function runif(1, min = 0, max = 4)). In my opinion, this is the best way to deal with the uncertainty as to which values are actually behind "4 oder weniger". To get a random value for each row, I needed to use the rowwise-function.

### Exercise 3c) 

**Task:** Calculate the total number of babies born in Munich in 2022
    and 2021. Which year had the bigger baby-boom?
    
```{r}

births_total_22 <- sum(names22$Anzahl)

births_total_21 <- sum(names21$Anzahl)

data.frame(Jahr = c("2021", "2022"), Anzahl = c(births_total_21, births_total_22)) %>% 
  slice_max(Anzahl)

```

**Answer:** Under the assumptions of 2b) **2021** had the bigger baby boom. 

### Exercise 3d) 

**Task:** Add a new column `year` to both datasets which holds the correct
    year for each.

```{r}
names21 <- names21 %>% 
  mutate(year = 2021)

names22 <- names22 %>% 
  mutate(year = 2022)
```

### Exercise 3e) 

**Task:** Combine both datasets into one using `bind_rows()`.

```{r}
names_total <- bind_rows(names21, names22)
```

### Exercise 3f)

**Task:** Combine the counts for same names to determine the most popular
    names across both years. Print out the top 10 names in a nicely
    formatted table for both years. Include a table caption.
    
```{r}
names_n <- names_total %>% 
  group_by(Vorname) %>%
  summarise(Häufigkeit = sum(Anzahl)) %>% 
  slice_max(order_by = Häufigkeit, n = 10)

knitr::kable(names_n, format = "html", caption = "Most popular names in Munich across 2022 and 2021") %>% 
  kableExtra::kable_styling()
```


## Exercise 4: Chat GPT + apply (3 points)

For this task: Specifically use ChatGPT to solve the task and submit
your prompts in addition to the solution

### Exercise 4a)

**Task:** The code below does not work because the wrong apply function has
    been used. Find out which apply function would be correct and why it
    did not work. Correct the code. Also calculate the rowwise means.

#### Create a sample data frame

`tax_data \<- data.frame( Name = c("Munich GmbH", "ABC Inc.", "Backpacks
1980", "Bavarian Circus"), Tax_2019 = c(5000, 4000, 6000, 3500),
Tax_2020 = c(4800, 4200, 5800, 3700), Tax_2021 = c(5200, 3800, 5900,
3400) )`

**Prompt for ChatGPT:** *I tried to create dataframe and assign it to an object in R. But unfortunately, I got this error: `Error: unexpected '\\' in "tax_data \"` Can you help me with that?*

**Corrected code:** 
```{r}
tax_data <- data.frame(
  Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"),
  Tax_2019 = c(5000, 4000, 6000, 3500),
  Tax_2020 = c(4800, 4200, 5800, 3700),
  Tax_2021 = c(5200, 3800, 5900, 3400)
)
```

#### Calculate column-wise means

`column_means \<- lapply(tax_data\[, -1\], 2, mean)`

**Prompt 1 for ChatGPT:** *In the created dataframe "tax_data" I want to calculate the means of each column. For this I created this code: `column_means \<- lapply(tax_data\[, -1\], 2, mean)`. Unfortunately, this expression doesn't give me the means of each columns. Instead I get this error message: Error: unexpected '\\' in "column_means \". Can you help me to find a solution? I'm looking for a solution with another apply-function.*

**Corrected code:** 
```{r}
column_means <- sapply(tax_data[, -1], mean)
```

**Prompt 2 for ChatGPT:** *Can you explain to me why my solution with the lapply-function didn't work? And why it was necessary to use the sapply-function?*

**Information I got from ChatGPT's response:** In the original code, lapply() was used with an incorrect syntax. The correct syntax would have been: `column_means <- lapply(tax_data[, -1], mean)`. However, using lapply() would have returned a list of means, rather than a vector. So, to obtain the means of each column as a vector, it is more appropriate to use sapply() instead of lapply(). 

#### Rowwise means

**Prompt 3 for ChatGPT:** Can you tell me how I can calculate the means of every row in the dataframe "tax_data" using an apply-function?

**Code:**
```{r}
row_means <- apply(tax_data[, -1], 1, mean)
```

### Exercise 4b)

**Task:** Using ChatGPT try to understand what the rapply() function does.
    Create an easy example with mock data where the function is used and
    explain it in your words.
    
**Prompt for ChatGPT:** Can you explain to me what the rapply-function in R does? 

**Creating mock data + application example:**
```{r}
test_list <- list(a = 1:3, b = list(c = 4:6, d = 7:9))

result <- rapply(test_list, function(x) x * 2)
```

**Explanation:** As all apply-functions, rapply() can be used to apply a function on all elements of a vector, data structure, etc. The rapply-function is particularly useful when working with nested data structures like lists in lists (like the example in the mock data). If my goal would be to apply a function on every number in the object `test_list` the rapply-function would be the best way to go as it applies the function on all levels of nesting. In the example above, the rapply-function makes it possible to multiply each element of the nested list with *2*. 
